---
title: "Bayesian Flexible Modelling of Spatially Resolved Transcriptomic Data"
author: "Arhit Chakrabarti"
output: 
    github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tutorial

We use the function *generate_data* to simulate matrix-variate spatial data. Sourcing the ``**BayesSRT_func.R**" file sources all necessary functions. We first simulate data from a matrix-variate normal distribution with 100 randomly generated spatial locations from $Uni(0,1)$ and 20 rows (genes). The true spatial covariance matrix is generated from a Matern kernel with smoothness $0.5$, range $1.5$ and marginal variance $1$. The true scale matrix $\Psi$ is generated from an AR structure i.e., with $\Psi_{i,j}= \rho^{|i - j|}$ where $\rho = 0.5$.  

```{r data_generation}
# First source the function BayesSRT_func.R file, which loads all the relevant 
# R functions (both custom functions and inbuilt functions)
suppressPackageStartupMessages(source("BayesSRT_func.R"))
datum <- generate_data(n_locs = 100, num_reps = 20, range = 1.5)
```

## Running the proposed MCMC algorithm

We run our proposed Blocked Gibbs sampler for 1000 iterations and consider first 500 samples as burn-in. We do not consider any thinning of the samples. We plot the traceplot of log-likelihood for each of the posterior samples from our sampler and the corresponding auto-correlation plot. We overlay the traceplot of log-likelihood with the true value of the log-likelihood.

```{r message=FALSE, warning=FALSE, results="hide"}
Blocked.MCMC = posterior_est(data = datum, n_iterations = 1000, burn = 500, thin = 1)
```

```{r gridextra, message=FALSE, warning=FALSE, results="hide", echo=FALSE}
if(!require(gridExtra)) install.packages("gridExtra"); suppressPackageStartupMessages(library(gridExtra))
```

```{r plots, echo=FALSE}
grid.arrange(Blocked.MCMC$LL.plot, Blocked.MCMC$ACF.plot, nrow = 2)
```

