---
title: "Bayesian Flexible Modelling of Spatially Resolved Transcriptomic Data"
output: 
    github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Single-Sample simulation

In this repository, we have a set of functions to simulate data from a single matrix-variate normal distribution and run our proposed Blocked Gibbs Sampling algorithm. We use the function **generate_data** to simulate matrix-variate spatial data. Sourcing the **BayesSRT_func.R** file sources all necessary functions.

For the purpose of illustration, we first simulate data from a matrix-variate normal distribution with 200 randomly generated spatial locations from the distribution, $Uni(0,1)$ and 20 rows (corresponding to genes in spatial transcriptomic data). The true spatial covariance matrix is generated from a Matern kernel with smoothness $0.5$, range $1.5$ and marginal variance $1$. 
The true scale matrix $\Psi$ is generated from an AR structure i.e., with $\Psi_{i,j}= \rho^{|i - j|}$ where $\rho = 0.5$.  

```{r data_generation}
# First source the function BayesSRT_func.R file, which loads all the relevant 
# R functions (both custom functions and inbuilt functions)
suppressPackageStartupMessages(source("BayesSRT_func.R"))
datum <- generate_data(n_locs = 200, num_reps = 20, range = 1.5)
```

## Running the proposed MCMC algorithm

We run our proposed Blocked Gibbs sampler for 2000 iterations and consider first 500 samples as burn-in. We do not consider any thinning of the samples. We plot the traceplot of log-likelihood for each of the posterior samples from our sampler and the corresponding auto-correlation plot. We overlay the traceplot of log-likelihood with the true value of the log-likelihood.

```{r message=FALSE, warning=FALSE, results="hide"}
Blocked.MCMC = posterior_est(data = datum, n_iterations = 2000, burn = 500, thin = 1)
```

```{r gridextra, message=FALSE, warning=FALSE, results="hide", echo=FALSE}
if(!require(gridExtra)) install.packages("gridExtra"); suppressPackageStartupMessages(library(gridExtra))
```

```{r plots, echo=FALSE}
grid.arrange(Blocked.MCMC$LL.plot, Blocked.MCMC$ACF.plot, nrow = 2)
```

## Multi-Sample simulation

We also provide the code for performing simulations for the multi-sample case. The associated R code is **simulation_replicates.R**. For simplicity, we are not reproducing the simulations in the README. 

## Real Data Analysis

The STARmap data used for the real data analysis can be found in the folder **STARmap**. The folder **Real Data Application** has the code **DataGeneration.R** which produces the analysis data from the STARmap datasets, which are also included in the same folder. The R file **BGS.R** runs the proposed Blocked Gibbs Sampler for the STARmap data. **LL_Correlation.R** file calculates the posterior log-likelihood values using the output from the **BGS.R** file. The **Analysis.R** file has the relevant codes used for generating all the exploratory plots and analysis relevant to the multi-sample real data analysis section.